Parallel Tractability of Ontology\\ Materialization: Technique and Practice

### 2018-06-25 parallel-tractability-J-v9/parallelTractabilityDatalog.tex
comma added, phrasing question

We now analyse the parallel complexity of Algorithm~$\mathsf{A}_{\text{opt}}$.
We use the same symbols $e$, $v$ and $r$ that are used for analyzing the complexity of Algorithm~$\mathsf{A}_{\text{bsc}}$
(see Section~\ref{sec:alg-bsc}). Similarly to the analysis for Algorithm~$\mathsf{A}_{\text{bsc}}$, it can be checked that
the number of processors is polynomial in $|\textbf{I}|$, i.e., $r(|\textbf{I}|e)^v$.
We now consider the computing time of Algorithm~$\mathsf{A}_{\text{opt}}$. It
is obvious that \ref{alg3:addFacts} and
\ref{alg3:halt} occupy constant time (denoted by $c_1$).
Algorithm~$\mathsf{A}_{\text{opt}}$ also works on the parallel condition.
That is each processor $q$ corresponds to a rule instantiation, e.g., $B_1,\ldots,B_n\rightarrow H$,
and processor $q$ can access the locations of nodes $B_1,\ldots,B_n, H$ in constant time.
In order to compute $S_{\!\textit{\tiny rch}}$, processor $q$ should check the derivability
of nodes $B_1,\ldots,B_n, H$. If only node $B_i$ ($i\in\{1,\ldots n\}$) and node $H$
are still not derived, then processor $q$ creates a pair $(B_i, H)$ in $S_{\!\textit{\tiny rch}}$.
The other elements in $S_{\!\textit{\tiny rch}}$ are similarly computed.
It is obvious that the above procedure occupies constant time as well on the parallel condition.
Thus, in \ref{alg3:updateG}, the phases of
computing $S_{\!\textit{\tiny rch}}$ and updating $\mathcal{G}$ also take constant time (denoted by $c_2$).
Recall that the number of constants is at most $|\textbf{I}|e$ (see Section~\ref{sec:alg-bsc}).
Let $w$ and $p$ represent the maximum arity of any predicate and the number of predicates, respectively.
We have that the number of all possible substitutions of atoms is at most $p(|\textbf{I}|e)^w$.
The relation $S_{\!\textit{\tiny rch}}$ consists of pairs of the atoms. Thus, the size of $S_{\!\textit{\tiny rch}}$
is bounded by $p^2(|\textbf{I}|e)^{2w}$. It can be checked that the computing time of the \texttt{NC} algorithm for computing the transitive
closure of $S_{\!\textit{\tiny rch}}$ is poly-logarithmic in the size of $\textbf{I}$, more precisely,
it is upper bounded by $\log^2(p^2(|\textbf{I}|e)^{2w})$ ($=4\log^2(p)+4w^2\log^2(|\textbf{I}|e)+8w\log(p)\log(|\textbf{I}|e)$).
We now have that the total computing time of
Algorithm~$\mathsf{A}_{\text{opt}}$ is $c_1+\ell c_3 + \ell
c_4\log(|\textbf{I}|e) +\ell c_5\log^2(|\textbf{I}|e)$
where $c_3=c_2+4\log^2(p)$, $c_4=8w\log(p)$, $c_5=4w^2$ and $\ell$ denotes the number of iterations of \ref{alg3:updateG}.

>>>>

We now analyse the parallel complexity of Algorithm~$\mathsf{A}_{\text{opt}}$.
We use the same symbols $e$, $v$ and $r$ that are used for analyzing the complexity of Algorithm~$\mathsf{A}_{\text{bsc}}$
(see Section~\ref{sec:alg-bsc}). Similarly to the analysis for Algorithm~$\mathsf{A}_{\text{bsc}}$, it can be checked that
the number of processors is polynomial in $|\textbf{I}|$, i.e., $r(|\textbf{I}|e)^v$.
We now consider the computing time of Algorithm~$\mathsf{A}_{\text{opt}}$. It
is obvious that \ref{alg3:addFacts} and
\ref{alg3:halt} occupy constant time (denoted by $c_1$).
Algorithm~$\mathsf{A}_{\text{opt}}$ also works on the parallel condition.
That is, each processor $q$ corresponds to a rule instantiation, e.g., $B_1,\ldots,B_n\rightarrow H$,
and processor $q$ can access the locations of nodes $B_1,\ldots,B_n, H$ in constant time.
In order to compute $S_{\!\textit{\tiny rch}}$, processor $q$ should check the derivability
of nodes $B_1,\ldots,B_n, H$. If only node $B_i$ ($i\in\{1,\ldots n\}$) and node $H$
are still not derived, then processor $q$ creates a pair $(B_i, H)$ in $S_{\!\textit{\tiny rch}}$.
The other elements in $S_{\!\textit{\tiny rch}}$ are similarly computed.
It is obvious that the above procedure requires constant time as well on the parallel condition.
Thus, in \ref{alg3:updateG}, the phases of
computing $S_{\!\textit{\tiny rch}}$ and updating $\mathcal{G}$ also take constant time (denoted by $c_2$).
Recall that the number of constants is at most $|\textbf{I}|e$ (see Section~\ref{sec:alg-bsc}).
Let $w$ and $p$ represent the maximum arity of any predicate and the number of predicates, respectively.
We have that the number of all possible substitutions of atoms is at most $p(|\textbf{I}|e)^w$.
The relation $S_{\!\textit{\tiny rch}}$ consists of pairs of the atoms. Thus, the size of $S_{\!\textit{\tiny rch}}$
is bounded by $p^2(|\textbf{I}|e)^{2w}$. It can be checked that the computing time of the \texttt{NC} algorithm for computing the transitive
closure of $S_{\!\textit{\tiny rch}}$ is poly-logarithmic in the size of $\textbf{I}$, more precisely,
it is upper bounded by $\log^2(p^2(|\textbf{I}|e)^{2w})$ ($=4\log^2(p)+4w^2\log^2(|\textbf{I}|e)+8w\log(p)\log(|\textbf{I}|e)$).
We now have that the total computing time of
Algorithm~$\mathsf{A}_{\text{opt}}$ is $c_1+\ell c_3 + \ell
c_4\log(|\textbf{I}|e) +\ell c_5\log^2(|\textbf{I}|e)$
where $c_3=c_2+4\log^2(p)$, $c_4=8w\log(p)$, $c_5=4w^2$ and $\ell$ denotes the number of iterations of \ref{alg3:updateG}.

### 2018-06-25 parallel-tractability-J-v9/response2.tex
comma added, phrasing question

Algorithm~$\mathsf{A}_{\text{opt}}$ also works on the parallel condition.
That is each processor $p$ corresponds to a rule instantiation, e.g., $B_1,\ldots,B_n\rightarrow H$,
and processor $p$ can access the locations of nodes $B_1,\ldots,B_n, H$ in constant time.
In order to compute $S_{\!\textit{\tiny rch}}$, processor $p$ should check the derivability
of nodes $B_1,\ldots,B_n, H$. If only node $B_i$ ($i\in\{1,\ldots n\}$) and node $H$
are still not derived, then processor $p$ creates a pair $(B_i, H)$ in $S_{\!\textit{\tiny rch}}$.
The other elements in $S_{\!\textit{\tiny rch}}$ are similarly computed.
On the parallel condition, the above procedure occupies constant time as well as checking
the applicability of a rule instantiation.
We have added the above content to make the claim that computation of $S_\text{rch}$ costs
constant time clearer.
The procedure of updating $G$ is similar to that of Algorithm~$\mathsf{A}_{\text{bsc}}$.

>>>>

Algorithm~$\mathsf{A}_{\text{opt}}$ also works on the parallel condition.
That is, each processor $p$ corresponds to a rule instantiation, e.g., $B_1,\ldots,B_n\rightarrow H$,
and processor $p$ can access the locations of nodes $B_1,\ldots,B_n, H$ in constant time.
In order to compute $S_{\!\textit{\tiny rch}}$, processor $p$ should check the derivability
of nodes $B_1,\ldots,B_n, H$. If only node $B_i$ ($i\in\{1,\ldots n\}$) and node $H$
are still not derived, then processor $p$ creates a pair $(B_i, H)$ in $S_{\!\textit{\tiny rch}}$.
The other elements in $S_{\!\textit{\tiny rch}}$ are similarly computed.
Under the parallel condition, the above procedure uses constant time as well as checking
the applicability of a rule instantiation. \textcolor{red}{BG: I
  cannot parse the previous sentence, in particular the `as well as
  checking\ldots' part.}
We have added the above arguments \textcolor{red}{to the paper?} to
make the claim that the computation of $S_\text{rch}$ costs
constant time clearer.
The procedure of updating $G$ is similar to that of Algorithm~$\mathsf{A}_{\text{bsc}}$.
